### 2023-09-20 chapter_2/src/slides/slides.qmd

---
title: "The Shape and Color of Politics"
author:
  - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: true
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 2
    slide-level: 5
    bibliography: ../../../assets/references.bib   
    self-contained: true
    output-file: color-book.html
execute:
  echo: false
---

@@

[![1860s Ballots](../../assets/slides/imgs/ballots.avif){#fig-ballots}](https://slate.com/news-and-politics/2020/02/ballot-history-photo-essay.html)

@@

[![Democratic National Convention](../../assets/slides/imgs/1980_dem_convention.jpg){#fig-1980-dem height=100}](https://untappedcities.com/2021/08/12/1980-democratic-national-convention/)

[![Republican National Convention](../../assets/slides/imgs/1980_rep_convention.jpg){#fig-1980-rep height=100}](https://reuther.wayne.edu/node/9207)

@@

[![2000 Presidential Election map](../../assets/slides/imgs/2000_election_map.webp){#fig-2000-election-map width=200}](https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.smithsonianmag.com%2Fhistory%2Fwhen-republicans-were-blue-and-democrats-were-red-104176297%2F&psig=AOvVaw3mlqWWKalyCl0gzVJDCIV5&ust=1694829600914000&source=images&cd=vfe&opi=89978449&ved=0CA8QjRxqFwoTCJCNg7bCq4EDFQAAAAAdAAAAABAT)

@@

[![The public](../../assets/slides/imgs/division.webp){#fig-public}](https://www.brookings.edu/articles/what-is-a-brokered-convention-what-is-a-contested-convention/)

@@

```{mermaid}
%%| label: fig-zaller
%%| fig-cap: Memory model of info processing
%%| fig-width: 6.5
graph TD;
A[Sampled pre-existing info] -->|Stronger impacts when less pre-existing knowledge| C[Updated attitudes]
B[New info] --> C
```

@@

![Red condition](../study_1/assets/rep_yard_sign_round_1.png)
![Blue condition](../study_1/assets/dem_yard_sign_round_1.png)
![White condition](../study_1/assets/cont_yard_sign_round_1.png)

@@

![Red condition](../study_1/assets/rep_yard_sign_round_2.png)
![Blue condition](../study_1/assets/dem_yard_sign_round_2.png)
![White condition](../study_1/assets/cont_yard_sign_round_2_3.png)

@@

![Red condition](../study_1/assets/rep_yard_sign_round_3.png)
![Blue condition](../study_1/assets/dem_yard_sign_round_3.png)
![White condition](../study_1/assets/cont_yard_sign_round_2_3.png)

@@

![Example of treatment](../../assets/slides/imgs/treatment.gif){#fig-treatment-example}

>>>>

---
title: "The Shape and Color of Politics"
author:
  - name: Damon C. Roberts [{{< fa brands orcid size=xs >}}](https://orcid.org/0000-0002-4360-3675)
format:
  revealjs: 
    embed-resources: true
    theme: [dark, ../../assets/slides/custom.scss]
    transition: slide
    scrollable: true
    incremental: true
    slide-number: true
    toc: true
    toc-depth: 2
    slide-level: 5
    bibliography: ../../../assets/references.bib   
    self-contained: true
    output-file: color-book.html
execute:
  echo: false
  freeze: true
  cache: true
---

@@

[![1860s Ballots](../../assets/slides/imgs/ballots.avif){#fig-ballots width=6.5in}](https://slate.com/news-and-politics/2020/02/ballot-history-photo-essay.html)

@@

[![Democratic National Convention](../../assets/slides/imgs/1980_dem_convention.jpg){#fig-1980-dem width=6.5in}](https://untappedcities.com/2021/08/12/1980-democratic-national-convention/)

[![Republican National Convention](../../assets/slides/imgs/1980_rep_convention.jpg){#fig-1980-rep width=6.5in}](https://reuther.wayne.edu/node/9207)

@@

[![2000 Presidential Election map](../../assets/slides/imgs/2000_election_map.webp){#fig-2000-election-map width=6.5in}](https://www.google.com/url?sa=i&url=https%3A%2F%2Fwww.smithsonianmag.com%2Fhistory%2Fwhen-republicans-were-blue-and-democrats-were-red-104176297%2F&psig=AOvVaw3mlqWWKalyCl0gzVJDCIV5&ust=1694829600914000&source=images&cd=vfe&opi=89978449&ved=0CA8QjRxqFwoTCJCNg7bCq4EDFQAAAAAdAAAAABAT)

@@

[![The public](../../assets/slides/imgs/division.webp){#fig-public width=6.5in}](https://www.brookings.edu/articles/what-is-a-brokered-convention-what-is-a-contested-convention/)

@@

```{mermaid}
%%| label: fig-zaller
%%| fig-cap: Memory model of info processing
graph TD;
A[Sampled pre-existing info] -->|Stronger impacts when less pre-existing knowledge| C[Updated attitudes]
B[New info] --> C
```

@@

![Red condition](../study_1/assets/rep_yard_sign_round_1.png){#fig-red-1 width="6.5in"}

![Blue condition](../study_1/assets/dem_yard_sign_round_1.png){#fig-blue-1 width="6.5in"}

![White condition](../study_1/assets/cont_yard_sign_round_1.png){#fig-white-1 width="6.5in"}

@@

![Red condition](../study_1/assets/rep_yard_sign_round_2.png){#fig-red-2 width=6.5in}

![Blue condition](../study_1/assets/dem_yard_sign_round_2.png){#fig-blue-2 width=6.5in}

![White condition](../study_1/assets/cont_yard_sign_round_2_3.png){#fig-white-2 width=6.5in}

@@

![Red condition](../study_1/assets/rep_yard_sign_round_3.png){#fig-red-3 width=6.5in}

![Blue condition](../study_1/assets/dem_yard_sign_round_3.png){#fig-blue-3 width=6.5in}

![White condition](../study_1/assets/cont_yard_sign_round_2_3.png){#fig-white-3 width=6.5in}

@@

![Example of treatment](../../assets/slides/imgs/treatment.gif){#fig-treatment-example width=6.5in}

### 2023-09-20 grants/ddirg/src/roberts_proposal_summary.qmd

The use of the colors red and blue to distinguish the two dominant political parties in the United States started during the 2000 Presidential Election with the electoral maps [@elving_2014_npr]. Since then, the two political party's logos have changed rather dramatically in that now the Republican party uses the color red, and the Democratic party only uses the color blue. @fig-party-brand nicely illustrates this point. Some existing descriptive research demonstrates that these clear distinctions in color are not limited to party logos but that Democrats use more blue than Republicans. In contrast, Republicans use more red on their campaign's yard signs [@williams_et-al_2022_jomp].^[The first empirical chapter of this dissertation replicates this finding when exploring the rates of color usage among partisans in districts that have more or less electoral competition.] There are two primary questions for this dissertation and book project. The first question seeks to answer whether the colors red and blue carry affiliative information for voters to use as a heuristic. The second question asks whether this affiliative information that these colors may provide influence attitude expression and behavior, even in the face of more traditional substantive political information.

@@

This grant would be a significant help with the costs of data collection. While I have performed the data collection for the first of the three primary empirical chapters, the last two require several experiments that require both online surveys and in-person experiments. Therefore, financial assistance with recruiting sufficiently powered and representative samples, along with adequately paying these participants for participating in these studies, would be a tremendous help for the completion of this dissertation and will remove the need for me to redo a number of these studies when working on the book project after defending the dissertation.

>>>>

The use of the colors red and blue to distinguish the two dominant political parties in the United States started with the electoral maps displayed during the 2000 presidential election [@elving_2014_npr]. Since then, the two political party's logos have changed rather dramatically in that now the Republican party only uses the color red, and the Democratic party only uses the color blue. @fig-party-brand illustrates this point. Some existing descriptive research demonstrates these clear distinctions in color are not limited  to party logos, but that Democrats use more blue than Republicans. In contrast, Republicans use more red on their campaign's yard signs [@williams_et-al_2022_jomp].^[The first empirical chapter of this dissertation replicates and extends this finding by exploring the rates of color usage among partisans in districts that have more/less electoral competition.] There are two primary questions for this dissertation and book project. I address the first question by laying out a theory of the cognitive processes explaining how these associations have crystallized for the mass public. I then explore the second question by examining what the consequences are of these affiliations for political attitude expression and behavior, even in the face of more traditional substantive political information (e.g., policy preferences).

A primary contribution of this project comes in its development of a cognitive model of visual information processing. Existing work on the visual aspects of politics is a newer area of work and comes from several different academic disciplines. Work to date leaves us with a somewhat fragmented theory of how visual information affects mass attitudes and behavior. This project lays out a model that illustrates how visual information, such as color, may be a source of information in politics; visual information is often processed faster than more symbolic information (e.g., substantive political information), so the first-order information explaining an individual's attitude expression may be visual information, such as color. The empirical chapters of the project discuss how these things influence our understanding of vote choice and the value of yard signs to campaigns, and how these things inform the degree to which people engage in conversations about politics. The project also provides a typology of when persuasion can occur, and it explains how we can make inferences about the partisanship of a group of people based on the way they have curated their environment.

This grant would be a significant help with the costs of data collection. While I have performed the data collection for the first of the three primary empirical chapters, the last two incorporate several experiments that require online samples. An APSA grant would help me recruit sufficiently powered and representative samples (and adequately pay participants for participation). The support offered by such a grant would not only help me complete the dissertation in the short term, but would give me a head start on the book version of the project (by eliminating the need for me to re-do the studies with higher quality samples after defending my dissertation).
<!--
The use of the colors red and blue to distinguish the two dominant political parties in the United States started during the 2000 Presidential Election with the electoral maps [@elving_2014_npr]. Since then, the two political party's logos have changed rather dramatically in that now the Republican party uses the color red, and the Democratic party only uses the color blue. @fig-party-brand nicely illustrates this point. Some existing descriptive research demonstrates that these clear distinctions in color are not limited to party logos but that Democrats use more blue than Republicans. In contrast, Republicans use more red on their campaign's yard signs [@williams_et-al_2022_jomp].^[The first empirical chapter of this dissertation replicates this finding when exploring the rates of color usage among partisans in districts that have more or less electoral competition.] There are two primary questions for this dissertation and book project. The first question seeks to answer whether the colors red and blue carry affiliative information for voters to use as a heuristic. The second question asks whether this affiliative information that these colors may provide influence attitude expression and behavior, even in the face of more traditional substantive political information.

@@

This grant would be a significant help with the costs of data collection. While I have performed the data collection for the first of the three primary empirical chapters, the last two require several experiments that require both online surveys and in-person experiments. Therefore, financial assistance with recruiting sufficiently powered and representative samples, along with adequately paying these participants for participating in these studies, would be a tremendous help for the completion of this dissertation and will remove the need for me to redo a number of these studies when working on the book project after defending the dissertation. 
-->
